### 2010-12-15 awareness.tex
Refactored paper so all sections are seperate documents

In this section, we give the explicit semantics to basic awareness logic,
originally presented in {\cite{fagin_belief_1987}}, with an additional axiom.
 One difference in our presentation is the inclusion of a novel letter and a
superficial change in notation.  This work is inspired by the developments in
van Benthem and Vel\'azquez-Quesada {\cite{van_benthem_inference_2009}}.

\begin{definition}
  Let $\Phi$ be a set of letters and define the language $\mathcal{L}_A
  (\Phi)$ as:
  \[ \phi \hspace{1em} : : = \hspace{1em} p \in \Phi \hspace{1em} |
     \hspace{1em} \circlearrowleft \hspace{1em} | \hspace{1em} \bot
     \hspace{1em} | \hspace{1em} \phi \rightarrow \psi \hspace{1em} |
     \hspace{1em} \Box \phi \hspace{1em} | \hspace{1em} A : \phi \hspace{1em}
  \]
\end{definition}

We define the other connectives and modal operators as usual for classical
modal logics.

\begin{definition}
  \label{awarenessmodels}An {\tmstrong{awareness model}} $\mathbbm{M}= \langle
  W, V, R, \mathcal{A} \rangle$ is a Kripke model with a valuation $V : \Phi
  \cup \{\circlearrowleft\} \rightarrow 2^W$ and an awareness function
  $\mathcal{A} : W \rightarrow 2^{\mathcal{L}_A (\Phi)}$.  We will usually
  denote $\mathcal{A} (w)$ as $\mathcal{A}_w$.

  Define the semantic entailment relation $\vDash$ where atomic propositions,
  logical connectives and modality are interpreted as usual.  We employ the
  following novel definition:
  \begin{eqnarray*}
    \mathbbm{M}, w \vDash A : \phi & \text{iff} & \phi \in \mathcal{A_w}
  \end{eqnarray*}
\end{definition}

The following definition distinguishes us from previously investigations of
awareness logic:

\begin{definition}
  Define the ``\tmtextbf{theory of a model} $\mathbbm{M}$'' as:
  \[ \tmop{Th} (\mathbbm{M}) \assign \{\phi \hspace{1em} | \hspace{1em}
     \mathbbm{M}, v \vDash \phi \text{ for all } v \in W\} \]
  We define the following properties that a model $\mathbbm{M}$ may posses:

  \begin{descriptiondash}
    \item[CSQ] $\mathbbm{M}, w \vDash \Box \phi \text{ iff } \tmop{Th}
    (\mathbbm{M}) \cup \mathcal{A}_w \vdash \phi$

    \item[SND] $\mathbbm{M}, w \vDash \circlearrowleft \text{ iff }
    \mathbbm{M}, w \vDash \phi$ for all $\phi \in \mathcal{A}_w$
  \end{descriptiondash}

  Here $\vdash$ is any logical consequence relation which is sound for
  $\vDash$, making true \tmtextbf{modus ponens} and \tmtextbf{reflection} (ie,
  $\Gamma \vdash \phi$ if $\phi \in \Gamma$)
\end{definition}

Neither \tmtextbf{CSQ} nor \tmtextbf{SND} correspond to simple modally
definable properties.

Intuitively, the \tmtextbf{CSQ} asserts that an agent believes a formula
$\phi$ if and only if it follows logically from their knowledge base and their
background knowledge, represented by $\tmop{Th} (\mathbbm{M})$. \
\tmtextbf{CSQ} is Vincent Hendrik's principle of logical omniscience mentioned
in {\S}\ref{intro}.

\tmtextbf{SND} asserts that $\circlearrowleft$ corresponds to the agent's
knowledge base being sound. A sound knowledge base will only render true
conclusions when used in deductions.  If one considers ``the existance of a
sound deduction'' sufficient for a kind of knowledge, then $\circlearrowleft$
is suitable for investigating this notion of epistemology.  That is, if
$\mathbbm{M}, w \vDash \circlearrowleft \wedge \Box \phi$, then we may
naturally say that the agent (implicitly) {\tmem{knows}} $\phi$ on the basis
of $\mathcal{A}_w$, rather than merely {\tmem{believing}} $\phi$ on the basis
of $\mathcal{A}_w$.

The logic of models making true \tmtextbf{CSQ} and \tmtextbf{SND} is provided
in Table \ref{logic1}.

\begin{table}[h]
  \begin{tabular}{l}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$\\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$\\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$\\
    $\vdash \Box(\phi \rightarrow \psi) \rightarrow \Box \phi \rightarrow \Box
    \psi$\\
    $\vdash A : \phi \rightarrow \Box \phi$\\
    $\vdash \circlearrowleft \rightarrow \Box \phi \rightarrow \phi$\\
    \\
    \begin{tabular}{lll}
      $\frac{\vdash \phi \rightarrow \psi \hspace{4em} \vdash \phi}{\vdash
      \psi}$ & {\hspace{6em}} & $\frac{\vdash \phi}{\vdash \Box \phi}$
    \end{tabular}
  \end{tabular}
  \caption{\label{logic1}Awareness Logic for \tmtextbf{CSQ} and
  \tmtextbf{SND}}
\end{table}

\begin{theorem}
  \label{completeness1}Assuming an infinite store of proposition letters
  $\Phi$, the awareness logic presented is sound and weakly complete for
  awareness models making true \tmtextbf{CSQ} and \tmtextbf{SND}.
\end{theorem}

\begin{proof}
  Soundness is straightforward, so we will only address completeness.  Assume
  $\nvdash \psi$.  Consider the finite canonical model $\mathbbm{M}= \langle
  W, V, R, \mathcal{A} \rangle$ formed of maximally consistent sets of
  subformulae of $\phi$ (closed under single negations{\footnote{A discussion
  of this property may be found in {\cite[pg.
  243]{blackburn_modal_2001}}.}}), as per the finitary modal completeness
  proofs found in {\cite[chapter 4.8]{blackburn_modal_2001}} and
  {\cite[chapter 5]{boolos_logic_1995}}.  Typical presentations do not
  include awareness, however it is defined intuitively in this case:
  \[ \phi \in \mathcal{A}_w \Longleftrightarrow A : \phi \in w \]
  Since $\psi$ is a subformula of itself, we have $\mathbbm{M}, w \nvDash
  \psi$ for some world $w$ by a finitary {\tmem{Lindenbaum Lemma}} and
  {\tmem{Truth Theorem}}.  Moreover, it is straightforward to verify that
  $\mathbbm{M}$ makes true the following properties:
  \begin{enumeratenumeric}
    \item $W$ and $\mathcal{A}_v$ are finite, and $\psi \in \mathcal{A}_v$
    only if $\psi$ is a (possibly negated) subformula of $\phi$

    \item if $\mathbbm{M}, v \vDash A : \psi$ then $\mathbbm{M}, v \vDash \Box
    \psi$

    \item if $\mathbbm{M}, v \vDash \circlearrowleft$ then $v R v$
  \end{enumeratenumeric}

  We next produce a bisimular{\footnote{The notion of bisimulation is
  discussed at length in {\cite[chapter 2.2]{blackburn_modal_2001}}.  We
  note that the basic definition and all of the usual theorems (modal
  equivalence, Henessy Milner, etc.) may be generalized to awareness models in
  a straightforward fashion. }} model $\mathbbm{M}'$ which makes true (1), (2)
  and a stronger from of (3):

  {\hspace*{\fill}}\begin{tabular}{ll}
    \tmtextbf{3$'$.} & $\mathbbm{M}', v \vDash \circlearrowleft$ iff $v R v$
  \end{tabular}{\hspace*{\fill}}

  To this end define $\mathbbm{M}' \assign \langle W', V', R', \mathcal{A}'
  \rangle$ such that{\footnote{Throughout this article, we use $l$ and $r$ to
  denote the two canonical injections associated with the coproduct $W \uplus
  W$. We will use $v_l$ and $v_r$ as the shorthand for $l (v)$ and $r (v)$
  respectively.}}: {\hspace*{\fill}}
  \begin{itemizedot}
    \item $W' \assign W \uplus W$

    \item $V' (p) \assign \{v_l, v_r \hspace{1em} | \hspace{1em} v \in V
    (p)\}$

    \item $R' \assign \{(v_l, u_r), (v_r, u_l) \hspace{1em} | \hspace{1em} v R
    u\} \cup \{(v_l, v_l), (v_r, v_r) \hspace{1em} | \hspace{1em} \mathbbm{M},
    v \vDash \circlearrowleft\}$

    \item $\mathcal{A}' (v_l) \assign \mathcal{A}' (v_r) \assign \mathcal{A}
    (v)$
  \end{itemizedot}

  It is straightforward to verify that $\mathbbm{M}'$ makes true the desired
  properties.  Let $Z \assign \{(v, v_l), (v, v_r) \hspace{1em} |
  \hspace{1em} v \in W\}$.  then $Z$ is a bisimulation between $\mathbbm{M}$
  and $\mathbbm{M}'$.  Therefore we know there is some $w \in W$ such that
  $\mathbbm{M}', w_l \nvDash \psi$ and $\mathbbm{M}', w_r \nvDash \psi$.

  Finally we construct a model $\mathbbm{M}''$ which agrees with
  $\mathbbm{M}'$ for all subformulae of $\psi$, which makes true
  \tmtextbf{CSQ} and \tmtextbf{SND}.  Let $\Lambda$ be the set of proposition
  letters occuring in $\psi$.  We will make use members of $\Phi \backslash
  \Lambda$ as {\tmem{nominals}}, as per the tradition in hybrid logic.  This
  is possible since both $W'$ and $\Lambda$ are finite and $\Phi$ is infinite
  by hypothesis.  Let $\iota : W \uplus W \hookrightarrow \Phi \backslash
  \Lambda$ be an injection assigning a fresh nominal associated with each
  world in $\mathbbm{M}'$.  Now define $\mathbbm{M}'' \assign \langle W'',
  V'', R'', \mathcal{A}'' \rangle$, where:

  \begin{center}
    \begin{tabular}{lllll}
      $\bullet$ & $W'' \assign W'$ &  & $\bullet$ & $V'' (p) \assign \left\{
      \begin{array}{ll}
        V' (p) & p \in \Lambda\\
        \{v\} & p = \iota (v)\\
        \varnothing & o / w
      \end{array} \right.$\\
      $\bullet$ & $R'' \assign R'$ & {\hspace{3em}} & $\bullet$ &
      $\mathcal{A}''_v \assign \{\phi \hspace{1em} | \phi \in \mathcal{A}'_v
      \} \cup \{\neg \iota (u) \hspace{1em} | \hspace{1em} \neg v R' u\}$
    \end{tabular}
  \end{center}

  By induction we have that for every subformula $\phi$ of $\psi$ that
  $\mathbbm{M}', v \vDash \phi$ if and only if $\mathbbm{M}'', v \vDash \phi$,
  hence $\mathbbm{M}'', w_l \nvDash \psi$.

  All that is left to show is that $\mathbbm{M}''$ makes true \tmtextbf{CSQ}
  and \tmtextbf{SND}.  $\mathbbm{M}''$ has three further properties:
  \begin{enumerateroman}
    \item $\mathcal{A}''_v$ is finite for all worlds $v$

    \item  $v R' u$ if and only if $\mathbbm{M}'', u \vDash \bigwedge
    \mathcal{A}'' (v)$

    \item The logic presented in Table \ref{logic1} is sound for
    $\mathbbm{M}''$
  \end{enumerateroman}
  From (ii) and the fact that $\mathbbm{M}'$ makes true (3$'$), we have
  \tmtextbf{SND} for $\mathbbm{M}''$.

  All that is left is to demonstrate \tmtextbf{CSQ} for some sound logic.  We
  will use the logic in Table \ref{logic1} itself.  From (i), (ii), (iii),
  and the \tmtextit{deduction theorem} for modal logic, we have the following
  line of reasoning:

  \begin{align*}
    Th({\mathbbm{M}}''){\cup}\mathcal{A}''_v{\vdash}{\phi} &
    {\Longleftrightarrow}Th({\mathbbm{M}}''){\vdash}\bigwedge\mathcal{A}''_v{\rightarrow}{\phi}\\
    &
    {\Longleftrightarrow}\bigwedge\mathcal{A}''_v{\rightarrow}{\phi}{\in}Th({\mathbbm{M}}'')\\
    &
    {\Longleftrightarrow}{\mathbbm{M}}'',u{\vDash}\bigwedge\mathcal{A}''_v{\rightarrow}{\phi}\text{
    for all $u \in W''$}\\
    & {\Longleftrightarrow}\text{ for all $u \in W''$, if
    }{\mathbbm{M}}'',u{\vDash}\mathcal{A}''_v\text{ then
    }{\mathbbm{M}}'',u{\vDash}{\phi}\\
    & {\Longleftrightarrow}\text{ for all $u \in W''$, if }v R 'u \text{ then
    }{\mathbbm{M}}'',u{\vDash}{\phi}\\
    & {\Longleftrightarrow}{\mathbbm{M}}'',v{\models}\Box{\phi}
  \end{align*}
\end{proof}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: