### 2010-12-15 conclusions.tex
another section factored out

(FIXME:  Needs to be extended a lot!)

The logic presented in Table \ref{logic1} is a conservative extension of the
basic modal logic $K$, which means that its decision problem is
\tmtextsf{PSPACE} hard (a lower bound).  Our finitary completeness proof
establishes that its complexity is in \tmtextsf{EXP2-TIME} (an upper bound).

Worlds in epistemic logic might correspond to fantastic, pretend scenarios;
however, it may be desirable for some philosophers to model knowledge about
imaginary lands.  After all, many children ``know'' a lot about Tolkien's
{\tmem{middle earth}} or Rowling's {\tmem{Hogwarts}} (the school Harry Potter
attends).