### 2010-12-15 awareness.tex
Factored out first completeness theorem

In this section, we give the explicit semantics to basic awareness logic,
originally presented in {\cite{fagin_belief_1987}}, with an additional axiom.
 One difference in our presentation is the inclusion of a novel letter and a
superficial change in notation.  This work is inspired by the developments in
van Benthem and Vel\'azquez-Quesada {\cite{van_benthem_inference_2009}}.

\begin{definition}
  Let $\Phi$ be a set of letters and define the language $\mathcal{L}_A
  (\Phi)$ as:
  \[ \phi \hspace{1em} : : = \hspace{1em} p \in \Phi \hspace{1em} |
     \hspace{1em} \circlearrowleft \hspace{1em} | \hspace{1em} \bot
     \hspace{1em} | \hspace{1em} \phi \rightarrow \psi \hspace{1em} |
     \hspace{1em} \Box \phi \hspace{1em} | \hspace{1em} A : \phi \hspace{1em}
  \]
\end{definition}

We define the other connectives and modal operators as usual for classical
modal logics.

\begin{definition}
  \label{awarenessmodels}An {\tmstrong{awareness model}} $\mathbbm{M}= \langle
  W, V, R, \mathcal{A} \rangle$ is a Kripke model with a valuation $V : \Phi
  \cup \{\circlearrowleft\} \rightarrow 2^W$ and an awareness function
  $\mathcal{A} : W \rightarrow 2^{\mathcal{L}_A (\Phi)}$.  We will usually
  denote $\mathcal{A} (w)$ as $\mathcal{A}_w$.

  Define the semantic entailment relation $\vDash$ where atomic propositions,
  logical connectives and modality are interpreted as usual.  We employ the
  following novel definition:
  \begin{eqnarray*}
    \mathbbm{M}, w \vDash A : \phi & \text{iff} & \phi \in \mathcal{A_w}
  \end{eqnarray*}
\end{definition}

The following definition distinguishes us from previously investigations of
awareness logic:

\begin{definition}
  Define the ``\tmtextbf{theory of a model} $\mathbbm{M}$'' as:
  \[ \tmop{Th} (\mathbbm{M}) \assign \{\phi \hspace{1em} | \hspace{1em}
     \mathbbm{M}, v \vDash \phi \text{ for all } v \in W\} \]
  We define the following properties that a model $\mathbbm{M}$ may posses:

  \begin{descriptiondash}
    \item[CSQ] $\mathbbm{M}, w \vDash \Box \phi \text{ iff } \tmop{Th}
    (\mathbbm{M}) \cup \mathcal{A}_w \vdash \phi$

    \item[SND] $\mathbbm{M}, w \vDash \circlearrowleft \text{ iff }
    \mathbbm{M}, w \vDash \phi$ for all $\phi \in \mathcal{A}_w$
  \end{descriptiondash}

  Here $\vdash$ is any logical consequence relation which is sound for
  $\vDash$, making true \tmtextbf{modus ponens} and \tmtextbf{reflection} (ie,
  $\Gamma \vdash \phi$ if $\phi \in \Gamma$)
\end{definition}

Neither \tmtextbf{CSQ} nor \tmtextbf{SND} correspond to simple modally
definable properties.

Intuitively, the \tmtextbf{CSQ} asserts that an agent believes a formula
$\phi$ if and only if it follows logically from their knowledge base and their
background knowledge, represented by $\tmop{Th} (\mathbbm{M})$. 
\tmtextbf{CSQ} is Vincent Hendrik's principle of logical omniscience mentioned
in {\S}\ref{intro}.

\tmtextbf{SND} asserts that $\circlearrowleft$ corresponds to the agent's
knowledge base being sound. A sound knowledge base will only render true
conclusions when used in deductions.  If one considers ``the existence of a
sound deduction'' sufficient for a kind of knowledge, then $\circlearrowleft$
is suitable for investigating this notion of epistemology.  That is, if
$\mathbbm{M}, w \vDash \circlearrowleft \wedge \Box \phi$, then we may
naturally say that the agent (implicitly) {\tmem{knows}} $\phi$ on the basis
of $\mathcal{A}_w$, rather than merely {\tmem{believing}} $\phi$ on the basis
of $\mathcal{A}_w$.

The logic of models making true \tmtextbf{CSQ} and \tmtextbf{SND} is provided
in Table \ref{logic1}.

\begin{table}[h]
  \begin{tabular}{l}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$\\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$\\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$\\
    $\vdash \Box(\phi \rightarrow \psi) \rightarrow \Box \phi \rightarrow \Box
    \psi$\\
    $\vdash A : \phi \rightarrow \Box \phi$\\
    $\vdash \circlearrowleft \rightarrow \Box \phi \rightarrow \phi$\\
    \\
    \begin{tabular}{lll}
      $\frac{\vdash \phi \rightarrow \psi \hspace{4em} \vdash \phi}{\vdash
      \psi}$ & {\hspace{6em}} & $\frac{\vdash \phi}{\vdash \Box \phi}$
    \end{tabular}
  \end{tabular}
  \caption{\label{logic1}Awareness Logic for \tmtextbf{CSQ} and
  \tmtextbf{SND}}
\end{table}

\begin{theorem}
  \label{completeness1}Assuming an infinite store of proposition letters
  $\Phi$, the awareness logic presented is sound and weakly complete for
  awareness models making true \tmtextbf{CSQ} and \tmtextbf{SND}.
\end{theorem}

\input{completeness1.tex}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: