### 2010-12-15 hybrid_awareness.tex

The method of the previous completeness proof makes implicit use of concepts from
hybrid logic{\footnote{Hybrid logic was first presented in
{\cite{prior_revised_1969}} and later formally developed in
{\cite{bull_approach_1970}}.}}.  In this section we adapt hybrid logic and
show that it is a complete for reasoning about knowledge bases.  We note that
because of the increased expressive power of hybrid logic, our completeness
theorem is simpler.

\begin{definition}
  Let $\Phi$ be a set of letters and $\Psi$ a set of nominals, and define the
  language $\mathcal{L}_H (\Phi, \Psi)$ as:
  \[ \phi \hspace{1em} : : = \hspace{1em} p \in \Phi \hspace{1em} |
     \hspace{1em} i \in \Psi \hspace{1em} | \hspace{1em} \bot \hspace{1em} |
     \hspace{1em} \phi \rightarrow \psi \hspace{1em} | \hspace{1em} \Box \phi
     \hspace{1em} | \hspace{1em} A : \phi \hspace{1em} | \hspace{1em} \forall
     \phi \]
\end{definition}

Our approach in hybrid logic is to employ a universal modality along with
nominals.  This framework presents a logic where the agent may reason about
various labeled scenarios.  Scenarios may be multiply labeled or not labeled
at all.  From these intuitions we have the following definition:

\begin{definition}
  \label{hybridsemantics}Let a \tmtextbf{hybrid model} $\mathbbm{M}= \langle
  W, V, R, \mathcal{A}, \ell \rangle$ be an awareness model as in Definition
  \ref{awarenessmodels}, along with a partial labeling function $\ell : \Psi
  \nrightarrow W$

  The semantics for $\vDash$ are the same as in Definition
  \ref{awarenessmodels}, only we have
  \begin{eqnarray*}
    \mathbbm{M}, w \vDash \forall \phi & \text{iff} & \forall v \in
    W.\mathbbm{M}, v \vDash \phi\\
    \mathbbm{M}, w \vDash i & \text{iff} & \ell (i) = w
  \end{eqnarray*}
\end{definition}

The other connectives and operators are defined as usual, although we employ a
few special shorthands:

\begin{definition}
  We employ the following special abbreviations:
  \begin{eqnarray}
    \text{\tmtextup{@}}_i \phi \assign \forall (i \rightarrow \phi)
    \hspace{2em} & \exists \phi \assign \neg \forall \neg \phi \hspace{3em} &
    \circlearrowleft_i \assign \diamondsuit i \nonumber
  \end{eqnarray}
\end{definition}

This gives rise to a validity reflecting one of the axioms we saw in
{\S}\ref{awarenesslogic}:
\[ \vDash \circlearrowleft_i \rightarrow \Box \phi \rightarrow
   \text{\tmtextup{@}}_i \phi \]
This can be read as ``If the agent's knowledge base is sound at world $i$,
then if they can deduce something from their knowledge base, what they deduce
must be true at world $i$.'.  In a way, this may be considered as
{\tmem{relativising}} knowledge particular named situations.

The semantics in Definition \ref{hybridsemantics} obviates the \tmtextbf{SND}
principle we previously presented, since there is no explicit
$\circlearrowleft$ symbol in this setting. \tmtextbf{CSQ} still makes sense
without modification, however.  The following gives a logic for hybrid models
making true \tmtextbf{CSQ}:

\begin{table}[h]
  \begin{tabular}{ll}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$ & $\vdash \forall (\phi
    \rightarrow \psi) \rightarrow \forall \phi \rightarrow \forall \psi$\\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$ & $\vdash \forall
    \phi \rightarrow \phi$\\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$ & $\vdash \forall \phi \rightarrow
    \forall \forall \phi$\\
    $\vdash \Box(\phi \rightarrow \psi) \rightarrow \Box \phi \rightarrow \Box
    \psi$ & $\vdash \exists \phi \rightarrow \forall \exists \phi$\\
    $\vdash A : \phi \rightarrow \Box \phi$ & $\vdash \forall \phi \rightarrow
    \Box \phi$\\
    & $\vdash i \rightarrow \phi \rightarrow \text{\tmtextup{@}}_i \phi$\\
    & \\
    \begin{tabular}{llll}
      $\frac{\vdash \phi \rightarrow \psi \hspace{4em} \vdash \phi}{\vdash
      \psi}$ & {\hspace{6em}}$\frac{\vdash \phi}{\vdash \Box \phi}$ &
      {\hspace{6em}} & $\frac{\vdash \phi}{\vdash \forall \phi}$
    \end{tabular} & 
  \end{tabular}
  \caption{\label{logic2}Hybrid Awareness Logic for \tmtextbf{CSQ}}
\end{table}

\begin{theorem}
  \label{completeness2}Assuming an infinite store of nominals $\Psi$, the
  hybrid awareness logic presented is sound and weakly complete for all hybrid
  models making true \tmtextbf{CSQ}.
\end{theorem}
\input{completeness2.tex}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: