### 2010-12-15 paper.tex
fixing references

\documentclass{article}

\usepackage{hyperref}
 \hypersetup{
     colorlinks,%
     citecolor=black,%
     filecolor=black,%
     linkcolor=black,%
     urlcolor=black
 }

\usepackage{stmaryrd}
\usepackage{todonotes}

\usepackage{geometry,amsmath,amssymb,wasysym,enumerate,color}
\geometry{letterpaper}

%%%%%%%%%% Start TeXmacs macros
\newcommand{\mathbbm}{\mathbb}
\newcommand{\assign}{:=}
\newcommand{\nin}{\not\in}
\newcommand{\tmem}[1]{{\em #1\/}}
\newcommand{\tmop}[1]{\ensuremath{\operatorname{#1}}}
\newcommand{\tmstrong}[1]{\textbf{#1}}
\newcommand{\tmtextbf}[1]{{\bfseries{#1}}}
\newcommand{\tmtextit}[1]{{\itshape{#1}}}
\newcommand{\tmtextsf}[1]{{\sffamily{#1}}}
\newcommand{\tmtextup}[1]{{\upshape{#1}}}
\newenvironment{descriptioncompact}{\begin{description} }{\end{description}}
\newenvironment{descriptiondash}{\begin{description} }{\end{description}}
\newenvironment{enumeratenumeric}{\begin{enumerate}[1.] }{\end{enumerate}}
\newenvironment{enumerateroman}{\begin{enumerate}[i.] }{\end{enumerate}}
\newenvironment{itemizedot}{\begin{itemize} \renewcommand{\labelitemi}{$\bullet$}\renewcommand{\labelitemii}{$\bullet$}\renewcommand{\labelitemiii}{$\bullet$}\renewcommand{\labelitemiv}{$\bullet$}}{\end{itemize}}
\newenvironment{proof}{\noindent\textbf{Proof\ }}{\hspace*{\fill}$\Box$\medskip}
\definecolor{grey}{rgb}{0.75,0.75,0.75}
\definecolor{orange}{rgb}{1.0,0.5,0.5}
\definecolor{brown}{rgb}{0.5,0.25,0.0}
\definecolor{pink}{rgb}{1.0,0.5,0.5}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
%%%%%%%%%% End TeXmacs macros

\begin{document}

\title{Epistemic Logics For Knowledge Bases}\author{Matthew P.
Wampler-Doty}\date{}\maketitle

\begin{abstract}
  In {\cite{benthem_reflections_1991}}, Johan van Benthem
  proposed a research program for logic of \tmtextit{explicit knowledge}. 
  Two efforts have since emerged with proposals for logic of explicit
  knowledge: the \tmtextit{Dynamics of Awareness}
  {\cite{benthem_inference_2009}} and \tmtextit{Justification Logic}
  {\cite{artemov_introducing_2005}}.  The purpose of this paper is to
  repurpose various epistemic logics from these previous efforts for reasoning
  about logical closures over knowledge bases.  The results in this paper are
  based on previous work carried out in
  {\cite{wampler-doty_evidentialist_2010}}.
\end{abstract}

\listoftodos

\section{Introduction}
\label{intro}
\input{introduction.tex}

\section{Basic Explicit Logics of Knowledge Bases}

\subsection{Awareness Logic}
\label{awarenesslogic}
\input{awareness.tex}

\subsection{Hybrid Awareness Logic}
\label{hybrid_awareness}
\input{hybrid_awareness.tex}

\section{Logics of Multiple Knowledge Bases}

So far, we have only dealt with systems that reason over a single knowledge
base.  In this section, we present various logics for reasoning over multiple
knowledge bases.

\subsection{Simple Justification Logic}
\label{simple_justification}
\input{simple_justification.tex}

\subsection{Lattice Justification Logic}
\label{lattice_justification}
\input{lattice_justification.tex}

\subsection{Neighborhood Semantics}
\label{neighborhood_semantics}
\input{neighborhood_semantics.tex}

\subsection{\label{quantifying}Using Modalities to Quantify over Knowledge
Bases}
\label{modal-modal}
\input{modal-modal.tex}

\section{Conclusions \& Further Research}
\label{conclusions}
\input{conclusions.tex}

\bibliographystyle{plain}
\bibliography{zotero}

\end{document}