### 2010-12-15 neighborhood_semantics.tex
factored out completeness theorem

Neighborhood semantics were originally developed by Dana Scott and Richard
Montague in the early 1970s as a generalization of Kripke semantics \
{\cite{montague_universal_2008,scott_advice_1970}}.  In
{\cite{fagin_belief_1988}}, Halpern and Fagin adapted neighborhood semantics
for reasoning about epistemic agents without logical omniscience.  In this
section will demonstrate how these semantics may be modified so that
neighborhoods corresponds to the logical consequences of a different knowledge
bases.  This allows for using logics with neighborhood semantics for
reasoning about multiple knowledge bases.  A point of distinction of
neighborhood semantics from JL is that it may be possible that two distinct
knowledge bases cannot be merged by the agent.

\begin{definition}
  Let $\Phi$ be a set of letters and define the language $\mathcal{L}_N (\Phi)$ as:
  \[ \phi\ : : =\ p \in \Phi\ |
    \ \bot\ |\ \phi \rightarrow \psi
    \ |\ \Box \phi\ |\ A :
     \phi\ |\ K \phi \]
\end{definition}

Just as previous semantics, ``$\Box \phi$'' is intended to mean that the agent
has an argument for $\phi$ from some knowledge base.  A novelty we present
here is that ``$K \phi$'' means that the agent has an argument for $\phi$ from
some \tmtextit{sound} knowledge base.

\begin{definition}
  \label{neighborhoodmodels}A {\tmstrong{neighborhood model}} $\mathbbm{M}=
  \langle W, V, \mathcal{N}, \mathcal{A} \rangle$ has a neighborhood function
  $\mathcal{N} : W \rightarrow 2^{2^W}$ and a multi-awareness function
  $\mathcal{A} : W \rightarrow 2^{2^{\mathcal{L}_N (\Phi)}}$

  The semantics for $\vDash$ have the following modifications:
  \begin{eqnarray*}
    \mathbbm{M}, w \vDash \Box \phi & \text{iff} & \text{there exists a $U \in
    \mathcal{N}_w$ where $\mathbbm{M}, v \vDash \phi$ for all $v \in U$}\\
    \mathbbm{M}, w \vDash K \phi & \text{iff} & \text{there exists a $U \in
    \mathcal{N}_w$ where $w \in U$ and $\mathbbm{M}, v \vDash \phi$ for all $v
    \in U$}\\
    \mathbbm{M}, w \vDash A : \phi & \text{iff} & \phi \in \bigcup
    \mathcal{A_w}
  \end{eqnarray*}
\end{definition}

Here we modify our previous notion of \tmtextbf{CSQ} and \tmtextbf{SND} to
match how our semantics are intended for reasoning over multiple bases:

\begin{definition}
  The following defines properties a neighborhood model may make true:

  \begin{descriptiondash}
    \item[NCSQ] $\mathbbm{M}, w \vDash \Box \phi \text{ iff there exists a set
    $X \in \mathcal{A}_w$ such that } \tmop{Th} (\mathbbm{M}) \cup X \vdash
    \phi$

    \item[NSND] $\mathbbm{M}, w \vDash K \phi \text{ iff there exists a set $X
    \in \mathcal{A}_w$ such that $\mathbbm{M}, w \vDash X$ and } \tmop{Th}
    (\mathbbm{M}) \cup X \vdash \phi$
  \end{descriptiondash}

  as before, $\vdash$ is any sound logical consequence relation for $\vDash$
  with \tmtextbf{modus ponens} and \tmtextbf{reflection}
\end{definition}

\begin{table}[h]
  \begin{tabular}{l}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$\\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$\\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$\\
    $\vdash K \phi \rightarrow \Box \phi$\\
    $\vdash K \phi \rightarrow \phi$\\
    $\vdash A : \phi \rightarrow \Box \phi$\\
    \\
    \begin{tabular}{llll}
      $\frac{\vdash \phi \rightarrow \psi \hspace{4em} \vdash \phi}{\vdash
      \psi}$ & {\hspace{6em}}$\frac{\vdash \phi \rightarrow \psi}{\vdash \Box
      \phi \rightarrow \Box \psi}$ & {\hspace{6em}} & $\frac{\vdash \phi
      \rightarrow \psi}{\vdash K \phi \rightarrow K \psi}$
    \end{tabular}
  \end{tabular}
  \caption{\label{logic3}A Neighborhood Logic for \tmtextbf{NCSQ} and
  {\tmstrong{NSND}}}
\end{table}

\begin{theorem}
  Assuming an infinite store of letters $\Phi$, the logic in Table
  \ref{logic3} is sound and weakly complete for neighborhood semantics making
  true \tmtextbf{NCSQ} and \tmtextbf{NSND}.
\end{theorem}

\input{completeness5.tex}

It is not too hard to identify the fragment of this logic that governs
knowledge alone.  The fragment given in Table \ref{logic4} has special
significance to us, since it governs the notion of knowledge presented in
{\S}\ref{quantifying}.

\begin{table}[h]
  \begin{tabular}{l}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$\\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$\\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$\\
    $\vdash K \phi \rightarrow \phi$\\
    \\
    \begin{tabular}{llll}
      $\frac{\vdash \phi \rightarrow \psi \hspace{4em} \vdash \phi}{\vdash
      \psi}$ &  &  & $\frac{\vdash \phi \rightarrow \psi}{\vdash K \phi
      \rightarrow K \psi}$
    \end{tabular}
  \end{tabular}
  \caption{\label{logic4}A knowledge-only neighborhood logic for
  \tmtextbf{NCSQ} and {\tmstrong{NSND}}}
\end{table}

\begin{proposition}
  The knowledge-only logic presented governs the $\Box$-free fragment of the
  logic in Table \ref{logic3}
\end{proposition}

This concludes our analysis of knowledge bases using neighborhood
semantics.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: