### 2011-01-23 awareness0.tex

In this section, apply basic awareness logic to Russell's
epistemology.  As we described in \S\ref{mod-russell}, we will use
awareness to model \emph{acquaintance}, and modality to model \emph{knowledge by description}.

\begin{definition}
  Let $\Phi$ be a set of letters and define the language $\mathcal{L}_A
  (\Phi)$ as:
  \[ \phi \  : : = \  p \in \Phi \  |
     \  \circlearrowleft \  | \  \bot
     \  | \  \phi \rightarrow \psi \  |
     \  \Box \phi \  | \  A : \phi \ 
  \]
\end{definition}

We define the other connectives and modal operators as usual for classical
modal logics.

\begin{definition}
  \label{awarenessmodels0}An {\tmstrong{awareness model}} $\mathbbm{M}= \langle
  W, V, R, \mathcal{A} \rangle$ is a Kripke model with a valuation $V : \Phi
  \cup \{\circlearrowleft\} \rightarrow 2^W$ and an awareness function
  $\mathcal{A} : W \rightarrow 2^{\mathcal{L}_A (\Phi)}$.  We will usually
  denote $\mathcal{A} (w)$ as $\mathcal{A}_w$.

  Define the semantic entailment relation $\vDash$ according to
  \textbf{awareness logic} as defined in \cite{fagin_belief_1988},
  where $A : \phi$ denotes awareness of $\phi$. 
% where atomic propositions,
%   logical connectives and modality are interpreted as usual.  We employ the
%   following novel definition:
%   \begin{eqnarray*}
%     \mathbbm{M}, w \vDash A : \phi & \text{iff} & \phi \in \mathcal{A_w}
%   \end{eqnarray*}
\end{definition}

Define the following property, as sketched in \S\ref{mod-russell}:

\begin{definition}
  \begin{descriptiondash}
    \item[CSQ$_\vdash$] $\mathbbm{M}, w \vDash \Box \phi \text{ iff } \tmop{Th}
    (\mathbbm{M}) \cup \mathcal{A}_w \vdash \phi$
    % \item[SND] $\mathbbm{M}, w \vDash \circlearrowleft \text{ iff }
    % \mathbbm{M}, w \vDash \phi$ for all $\phi \in \mathcal{A}_w$
  \end{descriptiondash}

  Here $\vdash$ is any logical consequence relation which is sound for
  $\vDash$, making true \tmtextbf{modus ponens} and \tmtextbf{reflection} (ie,
  $\Gamma \vdash \phi$ if $\phi \in \Gamma$)
\end{definition}

% Neither \tmtextbf{CSQ} nor \tmtextbf{SND} correspond to simple modally
% definable properties.

The logic of models making true \tmtextbf{CSQ} % and \tmtextbf{SND} 
is provided in Table \ref{logic0}.

\begin{table}[h]
  \begin{tabular}{l}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$\\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$\\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$\\
    $\vdash \Box(\phi \rightarrow \psi) \rightarrow \Box \phi \rightarrow \Box
    \psi$\\
    $\vdash A : \phi \rightarrow \Box \phi$\\
    $\vdash \circlearrowleft \rightarrow \Box \phi \rightarrow \phi$\\
    \\
    \begin{tabular}{lll}
      $\frac{\vdash \phi \rightarrow \psi \hspace{4em} \vdash \phi}{\vdash
      \psi}$ & {\hspace{6em}} & $\frac{\vdash \phi}{\vdash \Box \phi}$
    \end{tabular}
  \end{tabular}
  \caption{\label{logic0}Awareness Logic for Acquaintance and
    Knowledge by Description}
\end{table}

\begin{theorem}
  \label{completeness0}Assuming an infinite store of proposition letters
  $\Phi$, the awareness logic presented is sound and weakly complete for
  awareness models making true \tmtextbf{CSQ} and \tmtextbf{SND}.
\end{theorem}

\input{completeness1.tex}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: