### 2011-01-27 awareness.tex
threw away completeness proof in favor of a more economical remark

In this section, we modify the framework we have previously given so that both
knowledge and belief may be modeled.  Russell provides some intuition
about what distinguishes knowledge from belief\cite[pg. 63]{russell_problems_1936}:
\begin{quote}
But are we to say that nothing is knowledge except what is validly deduced
from true premisses? Obviously we cannot say this. Such a definition is at
once too wide and too narrow. In the first place, it is too wide, because it is
not enough that our premisses should be \emph{true}, they must also be
known.
\end{quote}
Russell may be interpreted as taking immediate knowledge to be
\emph{primitive} in the above passage.  To this end, we propose a new
symbol: $\PP$, which is a kind of letter.  We will interpret it to mean ``All of propositions
the agent is explicitly aware of are articles of knowledge.'' Unless $\PP$ is
true, the agent can only be said to possess logical belief.

It may be that everything the agent is explicitly aware of is accidentally true,
and therefore all of her beliefs derivative beliefs are true.
However, in this case we take it that the agent has a form of
\emph{right opinion}.

As before, we make the minimal assumption that knowledge is \emph{true}.

\begin{definition}
   $\mathcal{L}_\PP
  (\Phi)$ extends $\mathcal{L}_A(\Phi)$ as follows:
  \[ \phi \  : : = \  p \in \Phi \  | \  \bot
     \  | \  \phi \rightarrow \psi \  |
     \  \Box \phi \  | \  A : \phi \ |\ \circlearrowleft 
  \]
\end{definition}

\begin{definition}
  \label{awarenessmodels}A {\tmstrong{$\PP$-awareness model}} $\mathbbm{M}= \langle
  W, V, R, \mathcal{A} \rangle$ is a Kripke model with a valuation $V : \Phi
  \cup \{\circlearrowleft\} \rightarrow 2^W$, which is otherwise the
  same as an ordinary awareness model.  The semantics are unmodified,
  except that $\mathbb{M},w \models \PP \iff w \in V(\PP)$.
\end{definition}

Aside from \textbf{CSQ$_\vdash$}, we will be concerned with the
following new property

\begin{definition}\label{snd1}
  \begin{descriptiondash}
    \item[SND] If $\mathbbm{M}, w \vDash \circlearrowleft \text{ then }
    \mathbbm{M}, w \vDash \phi$ for all $\phi \in \mathcal{A}_w$
  \end{descriptiondash}
\end{definition}

The logic for \textbf{CSQ$_\vdash$} and \textbf{SND} differs from the
logic given in \S\ref{awarenesslogic0} by a single axiom:

\begin{table}[h]
\begin{centering}
  \begin{tabular}{l}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$\\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$\\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$\\
    $\vdash \Box(\phi \rightarrow \psi) \rightarrow \Box \phi \rightarrow \Box
    \psi$\\
    $\vdash A : \phi \rightarrow \Box \phi$\\
    $\vdash \PP \to \Box \phi \rightarrow \phi$\\
    \\
    \begin{tabular}{lll}
      $\displaystyle \frac{\vdash \phi \rightarrow \psi \ \  \vdash \phi}{\vdash
      \psi}$ & {\ \ } & $\displaystyle \frac{\vdash \phi}{\vdash \Box \phi}$
    \end{tabular}
  \end{tabular}
  \caption{\label{logic1}Awareness Logic for For Knowledge And
    Derivative Belief}
\end{centering}
\end{table}

\begin{theorem}
  \label{completeness1}Assuming an infinite store of proposition letters
  $\Phi$, the awareness logic presented is sound and weakly complete for
  awareness models making true \tmtextbf{CSQ}$_{\vdash'}$ and
  \tmtextbf{SND} for a suitable logic $\vdash'$.
\end{theorem}
\begin{proof}
The proof proceeds exactly as the proof of Theorem
\ref{completeness0}; the only difference is that the canonical model
and modified model make true \textbf{SND} instead of being reflexive.
\end{proof}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: