### 2011-01-27 paper.tex
Some new boilerplate

\documentclass{article}

\usepackage{hyperref}
 \hypersetup{
     colorlinks,%
     citecolor=black,%
     filecolor=black,%
     linkcolor=black,%
     urlcolor=black
 }

\usepackage{stmaryrd}
\usepackage{todonotes}

\usepackage{geometry,amsmath,amssymb,wasysym,enumerate,color}
\geometry{letterpaper}

%%%%%%%%%% Start TeXmacs macros
\newcommand{\mathbbm}{\mathbb}
\newcommand{\assign}{:=}
\newcommand{\nin}{\not\in}
\newcommand{\tmem}[1]{{\em #1\/}}
\newcommand{\tmop}[1]{\ensuremath{\operatorname{#1}}}
\newcommand{\tmstrong}[1]{\textbf{#1}}
\newcommand{\tmtextbf}[1]{{\bfseries{#1}}}
\newcommand{\tmtextit}[1]{{\itshape{#1}}}
\newcommand{\tmtextsf}[1]{{\sffamily{#1}}}
\newcommand{\tmtextup}[1]{{\upshape{#1}}}
\newenvironment{descriptioncompact}{\begin{description} }{\end{description}}
\newenvironment{descriptiondash}{\begin{description} }{\end{description}}
\newenvironment{enumeratenumeric}{\begin{enumerate}[1.] }{\end{enumerate}}
\newenvironment{enumerateroman}{\begin{enumerate}[i.] }{\end{enumerate}}
\newenvironment{itemizedot}{\begin{itemize} \renewcommand{\labelitemi}{$\bullet$}\renewcommand{\labelitemii}{$\bullet$}\renewcommand{\labelitemiii}{$\bullet$}\renewcommand{\labelitemiv}{$\bullet$}}{\end{itemize}}
\newenvironment{proof}{\noindent\textbf{Proof\ }}{\hspace*{\fill}$\Box$\medskip}
\definecolor{grey}{rgb}{0.75,0.75,0.75}
\definecolor{orange}{rgb}{1.0,0.5,0.5}
\definecolor{brown}{rgb}{0.5,0.25,0.0}
\definecolor{pink}{rgb}{1.0,0.5,0.5}
\newtheorem{definition}{Definition}
\newtheorem{notation}{Notation}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
%%%%%%%%%% End TeXmacs macros

\newcommand{\PP}{\circlearrowleft}
\newcommand{\nmodels}{\nvDash}
\begin{document}

\title{Awareness and Russell's Epistemology}\author{Matthew P.
Wampler-Doty}\date{}\maketitle

\begin{abstract}
This paper presents a novel approach to formulating epistemic directly
in terms of logical inference.  Using awareness logic, $\Box_X \phi$ is
taken to literally mean ``it follows from formulae $X$ is explicitly aware of
that $\phi$''.  This is done by incorporating logic into the semantics
of awareness logic.  Our approach corresponds closely to the framework
for epistemology proposed by Bertrand
Russell\cite{russell_problems_1936}, and illustrates that
objections provided by Jaako Hintikka\cite{hintikka_knowledge_1969} to
the proposed interpretation of epistemic logic should be considered invalid.
\end{abstract}

\listoftodos
\section{Introduction}
\label{intro}
\input{introduction.tex}

\section{Philosophical Background}
\label{trad}
\input{foundationalism.tex}

\section{Basic Explicit Logics of Foundationalism}

\subsection{Acquaintance and Knowledge By Description}
\label{awarenesslogic0}
\input{awareness0.tex}

\subsection{Knowledge and Belief}
\label{awarenesslogic}
\input{awareness.tex}

\subsection{Hybrid Logic}
\label{hybrid_awareness}
\input{hybrid_awareness.tex}

\section{Logics of Multiple Awareness Sets}

So far, we have only dealt with systems for reasoning over a single knowledge
base.  In this section, we present various logics for reasoning over multiple
knowledge bases.

\subsection{Simple Justification Logic}
\label{simple_justification}
\input{simple_justification.tex}

% \subsection{Lattice Justification Logic}
% \label{lattice_justification}
% \input{lattice_justification.tex}

\subsection{Neighborhood Semantics}
\label{neighborhood_semantics}
\input{neighborhood_semantics.tex}

\subsection{\label{quantifying}Multi-Modal Logic}
\label{modal-modal}
\input{modal-modal.tex}

\section{Conclusions \& Further Research}
\label{conclusions}
\input{conclusions.tex}

\bibliographystyle{plain}
\bibliography{zotero}

\end{document}