### 2011-01-30 simple_justification.tex
Reworded introduction

Justification Logic (JL) was originally developed by Artemov as the
\tmtextit{Logic of Proofs} (LP) {\cite{artemov_logic_1994}}.  Its original
purpose was to provide a framework for reasoning about explicit provability in
Peano Arithmetic.  The first introduction of Justification Logic can be found
in {\cite{artemov_introducing_2005}}, where Artemov and Nogina propose LP as a
logic for reasoning about evidence.  In {\cite{fitting_logic_2005}}, Melvin
Fitting provided JL with Kripke model based semantics.

The principle innovation of LP/JL is to extend awareness logic, such that
awareness operations are now \tmtextit{proof terms}.  Informally, we say that
``a proof term $X$ witnesses proposition $\phi$'', denoted ``$X : \phi$'',
whenever $X$ represents a proof of $\phi$.  Proofs may have
{\tmstrong{multiple-conclusions}} in this system, so it is possible that $X :
\phi$ and $X : \psi$ can be true where $\phi \neq \psi$.  In Fitting's Kripke
semantics, $X : \phi$ means that $\phi$ is in awareness set corresponding to
$X$.  Proof terms are thought be operating in a \tmtextit{multi-conclusion}
proof system, so the same proof term may witness many different propositions.

The logics LP/JL include operators over proofs terms, so that new proof terms
may be assembled.  One operation of particular interest to us is
{\tmem{choice}}, denoted $\oplus$.  The expression ``$X \oplus Y : \phi$''
denotes that either $X$ or $Y$ are proofs witnessing $\phi$.  There are other
operations which correspond to \tmtextit{modus ponens} and
\tmtextit{proof-theoretic reflection}, however we do not consider these
operations here.

In this section we propose a stripped down, simplified form of JL suitable for reasoning
about derivative knowledge from awareness sets, which we call 
\tmtextit{simple justification logic} (SJL). Instead of reading $X :
\phi$ as ``$X$ is a proof term deriving $\phi$'', we will read this
expression as ``$X$ is acquainted with $\phi$''\footnote{This is
  inspired by Fitting's semantics for JL proposed in
  \cite{fitting_logic_2005}.}.  We read $\Box_X\phi$ as ``$X$ has
derivative knowledge that $\phi$''.
% However, instead of thinking of terms as representing proofs, we
% use them to represent knowledge bases. This is in the spirit of JL as a logic
% of evidence: we consider each knowledge base to be a corpus of evidence. \
% Each term ``names'' a different knowledge base at a particular world.  Terms
% may not refer to the same knowledge base at different worlds, just as in how
% in awareness logic the agent need not be aware of the same formulae at
% different worlds.
SJL reduces JL to only the {\tmem{choice}} operator, which is used as a 
mechanism for forming the union of awareness sets.

While the intuition behind SJL cannot be found in
Russell's writings, it may be found elsewhere in the epistemic logic
literature.  For instance in
\cite[pg. 67]{meyer_epistemic_1995}, Meyer et al. motivate logic for what they call
\emph{implicit knowledge in a group} with the following thought experiment:
\begin{quote}
[A]ssume $\phi$ to be `P = NP'. We furthermore assume that three
researchers 1, 2, 3 work on this long-standing and famous problem in
order to discover a proof of the theorem. Suppose that the actual
situation is such that $\phi$ follows from three lemmas: $\psi_1$,
$\psi_2$ and $\psi_3$. Researcher 1 has proved $\psi_1$ -- so he knows
$\psi_1$. Analogously for researchers 2 and 3 with respect to $\psi_2$
and $\psi_3$, respectively. \ldots If the researchers would interact
at a congress on theoretical computer science, they would be able to
gather their partial knowledge and thereby know (a proof of) $\phi$.
\end{quote}
This motivation works just as well for SJL. We model this situation by
employing the following symbolism:
there are three awareness sets $\mathcal{A}_1(w)$, $\mathcal{A}_2(w)$ and
$\mathcal{A}_3(w)$ which correspond to facts each researcher is
acquainted with.  Moreover, each lemma may be derived from its
respective base: $Th(\mathbb{M}) \cup \mathcal{A}_1(w) \vdash
\psi_1$, $Th(\mathbb{M}) \cup \mathcal{A}_1(w) \vdash
\psi_2$ and $Th(\mathbb{M}) \cup \mathcal{A}_1(w) \vdash
\psi_3$.  Finally, it holds that the three lemmas prove the theorem:
$Th(\mathbb{M}) \vdash \psi_1 \to \psi_2 \to \psi_3 \to \phi$.
Provided that our notion of logical consequence is \emph{monotonic},
and makes true modus ponens, it then follows that
$Th(\mathbb{M}) \cup \mathcal{A}_1(w) \cup \mathcal{A}_2(w) \cup
\mathcal{A}_3(w) \vdash \phi$.  In order to express this in the semantics, we represent $\mathcal{A}_1(w) \cup \mathcal{A}_2(w) \cup
\mathcal{A}_3(w)$ as $\mathcal{A}_{1 \oplus 2 \oplus 3}(w)$, and
$Th(\mathbb{M}) \cup \mathcal{A}_1(w) \cup \mathcal{A}_2(w) \cup
\mathcal{A}_3(w) \vdash \phi$ as $\mathbb{M}, w \models \Box_{1 \oplus 2 \oplus 3} \phi$.

\begin{definition}
  Let $\Pi$ be a set of primitive terms.  Define $\tau (\Pi)$ with the
  following grammar:
  \[ X \  : : = \  t \in \Pi \  | \  X  \oplus Y \]

  Let $\Phi$ be a set of letters and $\Pi$ a set of primitive terms, and
  define the language $\mathcal{L}_{\tmop{SJL}} (\Phi, \Pi)$ as:
  \[ \phi \  : : = \  p \in \Phi \  | \  \bot
     \  | \  \phi \rightarrow \psi \  |
     \  \Box_X \phi \  | \  X : \phi \]
  where $X \in \tau (\Pi)$
\end{definition}

\begin{definition}
  \label{justmodels}A {\tmstrong{simple justification model}} $\mathbbm{M}=
  \langle W, V, R, \mathcal{A} \rangle$ is a Kripke model with a valuation $V
  : \Phi  \rightarrow 2^W$, an indexed relation $R : \tau (\Pi) \rightarrow
  2^{W \times W}$, along with a modified awareness function
  $\mathcal{A} : \tau (\Pi) \times W
   \rightarrow 2^{\mathcal{L}_{\tmop{SJL}} (\Phi, \Pi)}$. 
  In practice we will denote $\mathcal{A} (X, w)$ by a curried shorthand, namely $\mathcal{A}_X (w)$.

  The semantics for $\vDash$ have the following modifications:
  \begin{eqnarray*}
    \mathbbm{M}, w \vDash \Box_X \phi & \iff & \text{for all $v \in W$
    where $w R_X v$ we have $\mathbbm{M}, v \vDash v$}\\
    \mathbbm{M}, w \vDash X : \phi & \iff & \phi \in \mathcal{A}_X (w)
  \end{eqnarray*}
\end{definition}

\begin{definition}
A Hilbert-style calculus $\vdash$ is said to be \textbf{SJL-suitable} for an SJL model
$\mathbb{M}$ if and only if it makes true modus ponens, it is
relatively sound for $\mathbb{M}$, and it is \textbf{monotonic}: if
$\Gamma \vdash \phi$ and $\Gamma \subseteq \Delta$, then $\Delta
\vdash \phi$.
\end{definition}

\begin{definition}
  The following defines properties a simple justification model may make true:

  \begin{descriptiondash}
    \item[$\textup{\textbf{JCSQ}}_\vdash$] $\mathbbm{M}, w \vDash \Box_X \phi \iff \tmop{Th}
    (\mathbbm{M}) \cup \mathcal{A}_X (W) \vdash \phi$
    \item[CHOICE] $\mathcal{A}_{X \oplus Y}(w) = \mathcal{A}_X (w) \cup
    \mathcal{A}_Y (w)$
  \end{descriptiondash}
\end{definition}

\begin{table}[h]
\begin{centering}
  \begin{tabular}{ll}
    $\vdash \phi \rightarrow \psi \rightarrow \phi$ & \\
    $\vdash (\phi \rightarrow \psi \rightarrow \chi) \rightarrow (\phi
    \rightarrow \psi) \rightarrow \phi \rightarrow \chi$ & \\
    $\vdash ((\phi \rightarrow \bot) \rightarrow (\psi \rightarrow \bot))
    \rightarrow \psi \rightarrow \phi$ & \\
    $\vdash \Box_X (\phi \rightarrow \psi) \rightarrow \Box_X \phi \rightarrow
    \Box_X \psi$ & \\
    $\vdash (X : \phi) \rightarrow \Box_X \phi$ & \\
    $\vdash \Box_X \phi \rightarrow \phi$ & \\
    $\vdash (X \oplus Y : \phi) \leftrightarrow (X : \phi \vee Y :
    \phi)$ & \\
    $\vdash \Box_X \phi \rightarrow \Box_{X \oplus Y} \phi$ & \\
    $\vdash \Box_{X \oplus X} \phi \rightarrow \Box_X \phi$ & $\vdash \Box_{(X
    \oplus X) \oplus Y} \phi \rightarrow \Box_{X \oplus Y} \phi$ \\
    $\vdash \Box_{X \oplus Y} \phi \rightarrow \Box_{Y \oplus X} \phi$ &
    $\vdash \Box_{(X \oplus Y) \oplus Z} \phi \rightarrow \Box_{(Y \oplus X)
    \oplus Z} \phi$\\
    $\vdash \Box_{(X \oplus Y) \oplus Z} \phi \rightarrow \Box_{X \oplus (Y
    \oplus Z)} \phi$ & \\
    & 
  \end{tabular}
\begin{tabular}{lll}
      $\displaystyle\frac{\vdash \phi \rightarrow \psi \ \  \vdash \phi}{\vdash
      \psi}$ & {\ \ } & $\displaystyle\frac{\vdash \phi}{\vdash \Box_X \phi}$
    \end{tabular}
  \caption{\label{logic5}Simple Justification Logic}
\end{centering}
\end{table}
We assert without proof that SJL is a conservative extension of the awareness logic in
{\S}\ref{awarenesslogic0}.

\begin{theorem}
  \label{completeness5}Assuming an infinite store of proposition letters
  $\Phi$, SJL (Table \ref{logic5}) is sound and weakly complete for reflexive simple justification models
  which make true $\textup{\tmtextbf{JCSQ}}_{\vdash'}$ for some
  SJL-suitable logic $\vdash'$ as well as
  \tmtextbf{CHOICE}
\end{theorem}
\input{completeness3}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: