### 2011-01-30 introduction.tex
Fixed reference to hendricks

Modern epistemic logic is commonly acknowledged to begin with
Hintikka's \emph{Knowledge and Belief}
\cite{hintikka_knowledge_1969}. While Hintikka began by considering
traditional topics in the theory of knowledge, epistemic logic
resulted in a bifurcation 20th century research into epistemology.  A number of
logicians and philosophers have acknowledged this rift, most notably
Vincent Hendricks\cite{hendricks_wheres_2006,hendricks_mainstream_2006}.

The purposes of this essay is to try to close the gap between these
traditions, by showing that awareness logics are suitable for
reasoning about a form of strong foundationalism proposed
by Bertrand Russell in \emph{The Problems of Philosophy}
\cite{russell_problems_1936}.  We begin by reviewing Russell's
theory, and contrasting it with modern perspectives.  In particular,
we illustrate that the modern concepts of \emph{explicit} and
\emph{implicit} knowledge, originally due to Levesque\footnote{While we will refer often to
  terms coined by Levesque in this essay, we will not use his
  proposed semantics.  We instead prefer variations on
  \emph{awareness logic}, originally developed by Fagin and Halpern
  \cite{fagin_belief_1988}.}\cite{levesque_logic_1984},  are reflect Russell's distinction between \emph{immediate} and \emph{derivative}
knowledge.  We propose a novel property of awareness models which
is inspired by Russell's epistemology.  The
remainder of this paper is then devoted to exploring how various awareness
logics can be designed to express this foundationalist property.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: